Hypertension and microbiome

### 2019-06-04 articletwo.Rmd
Not limited to twenty taxa
microbiome global renamed alpha

    maaslin.abund <- abundances(pseq.genus.core, transform = "compositional") %>%
        t %>%
        as.data.frame %>%
        select(-contains("virus")) %>%
        select(1:20)

    maaslin <- Maaslin2(maaslin.abund,
             maaslin.pheno,
             paste0("cache/maaslin-", now),
             transform = 'AST',
             fixed_effects = paste0(var.CL, collapse = ","), 
             cores = 10)

@@

```{r alphabeta definitions}
dset.alpha <- meta.merge.alphadiversity(pseq.species) %>%
    mutate(diversities_shannon = scale(diversities_shannon),
           MAP = oMAP,
           PULSEPRESSURE = oPULSEPRESSURE,
           SYSTM = oSYSTM,
           DIASM = oDIASM)

alphadiversity <- list(
    min = calculateglm(dset.alpha,
                       modelstring = "%s ~ BL_AGE + SEX  + diversities_shannon"),
    max = calculateglm(dset.alpha,
                       modelstring = "%s ~ BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09 + diversities_shannon"))

diversities <- mergediversities(alphadiversity, adonis.species)
```

@@

```{r difference in beta diversity}
merge(diversities$min %>% select(response, R2),
      diversities$max %>% select(response, R2), by = "response") %>%
    mutate(R2.diff = R2.y - R2.x) %>%
    kable
```

@@

Results for beta diversity

```{r table beta diversity}
adonis.species$max %>%
    data.table::rbindlist(id = "model_name") %>%
    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
    dplyr::select(response, R2, "Pr(>F)") %>%
    mutate(p.value = pub.p(`Pr(>F)`)) %>%
    kable
```

```{r save grob}
alpha.diversity.min <- plot.alpha(diversities$min, colorlimit = 0.6)
alpha.diversity.max <- plot.alpha(diversities$max, colorlimit = 0.6)

@@

abund.genus.bacteria <- colnames(abund.genus)[-1] %>% mygrep(".bacteria$", .)

@@

```{r pcoa axis analysis axis 1}
calculateglm(pcoa.df,
             modelstring = "%s ~ Axis.1 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
             filterstr = "Axis") %>% 
    select(response, term, estimate, std.error, p.value) %>%
    mutate_if(is.numeric, round, 3) %>%
    kable
```

Testing associations to blood pressure variables on PCoA axis 2

```{r pcoa axis analysis axis 2}
calculateglm(pcoa.df,
             modelstring = "%s ~ Axis.2 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
             filterstr = "Axis") %>% 
    select(response, term, estimate, std.error, p.value) %>%
    mutate_if(is.numeric, round, 3) %>%
    kable
```

Testing associations to blood pressure variables on PCoA axis 3

```{r pcoa axis analysis axis 3}
calculateglm(pcoa.df,
             modelstring = "%s ~ Axis.3 + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                     BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09",
             filterstr = "Axis") %>%
    select(response, term, estimate, std.error, p.value) %>%
    mutate_if(is.numeric, round, 3) %>%
    kable
```

```{r pcoa 3d}
theta <- 130
phi <- 10
pcoa3d.plot <- ggplot(data=pcoa.df, aes(x=Axis.1, y = Axis.2, z=Axis.3, color = HYPERTENSION)) +
  axes_3D(theta=theta, phi=phi) +
  stat_3D(theta=theta, phi=phi, size = 0.01, alpha=0.4) +
  labs_3D(theta=theta, phi=phi, 
          hjust=c(1,0,0), vjust=c(1.5,1,-.2),
          labs=c(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1]),
                 sprintf("PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2]),
                 sprintf("PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3]))) #+
#  theme_void()

ggsave(file = "cache/pcoa-species-3d.png", plot=pcoa3d.plot, height=4, width=7, units = "in", dpi = 300)

>>>>

    maaslin.abund <- abundances(pseq.genus.core, transform = "compositional") %>%
        t %>%
        as.data.frame %>%
        select(-contains("virus"))

    maaslin <- Maaslin2(maaslin.abund,
             maaslin.pheno,
             paste0("cache/maaslin-", now),
             transform = 'AST',
             fixed_effects = paste0(var.CL, collapse = ","), 
             cores = detectCores())

@@

```{r alphabeta definitions}
dset.alpha <- meta.merge.alphadiversity(pseq.species) %>%
    mutate(diversity_shannon = scale(diversity_shannon),
           MAP = oMAP,
           PULSEPRESSURE = oPULSEPRESSURE,
           SYSTM = oSYSTM,
           DIASM = oDIASM)

alphadiversity <- list(
    min = calculateglm(dset.alpha,
                       modelstring = sprintf("%%s ~ %s + diversity_shannon",
                                             paste0(var.CL.min, collapse = " + ")),
                       filterstr = "shannon"),
    max = calculateglm(dset.alpha,
                       modelstring = sprintf("%%s ~ %s + diversity_shannon",
                                             paste0(var.CL, collapse = " + ")),
                       filterstr = "shannon"))
```

@@

Difference of R2 between full and minimal models

```{r difference in beta diversity}
merge(diversities$min %>%
      select(response, R2, R2.p) %>%
      dplyr::rename(R2.min = R2, p.min = R2.p),
      diversities$max %>%
      select(response, R2, R2.p) %>%
      dplyr::rename(R2.full = R2, p.full = R2.p),
      by = "response") %>%
    mutate(R2.diff = R2.full - R2.min) %>%
    kable
```

@@

```{r save grob}
alpha.diversity.min <- plot.alpha(diversities$min, alphabreaks = seq(-1,0,0.2))
alpha.diversity.max <- plot.alpha(diversities$max, alphabreaks = seq(-1,0,0.2))

@@

abund.genus.bacteria <- abund.genus %>% select(-contains("virus")) %>% colnames

@@

Significant associations between BP-variables and first three PCoA axis

```{r pcoa axis analysis axis 1}
glm.pcoa <- lapply(c2l("Axis.1", "Axis.2", "Axis.3"), function(axis)
    calculateglm(pcoa.df,
                 modelstring = sprintf("%%s ~ %s + %s", axis, paste(var.CL, collapse = "+")),
                 filterstr = "Axis"))

do.call(rbind, glm.pcoa) %>% 
    select(response, term, estimate, std.error, p.value) %>%
    mutate_if(is.numeric, round, 3) %>%
    filter(p.value < 0.05) %>%
    arrange(p.value) %>%
    kable
```
